### 2018-02-17 Writing/Thesis/content/abstract.tex
Some more corrections (writing)

% Motivation
Die weltweit verfügbare Menge an Daten steigt immer weiter an, und doch bleibt der Anteil der tatsächlich analysierten Daten gering. Dieses Problem entsteht, da nicht genug Experten verfügbar sind, um diese Analyse auszuführen, und es für Laien schwierig ist, das richtige Verfahren auszuwählen. Besonders im Bereich der Klassifikation gibt es viele Algorithmen, deren Performanz jedoch von Datensatz zu Datensatz variieren kann - es gibt also keinen einen besten Algorithmus, der immer eine gute Wahl darstellt. Eine Antwort auf diese Problem ist automatisiertes Machine Learning. Daher werden in dieser Arbeit zwei verschiedene Ansätze, welche für einen neuen Datensatz ein Ranking von Klassifizierungsalgorithmen aufgrund von Eigenschaften des Datensatzes vorhersagt, vorgestellt und miteinander verglichen. Der erste der beiden Ansätze arbeitet auf Grundlage von Regressionsmodellen; ein Ranking wird hier konstruiert indem ein einzelnes Regressionsmodell für jeden Klassifizierungsalgorithmus einen Performanzwert vorhersagt, auf deren basierend dann das Ranking erstellt wird. Der zweite Ansatz beschäftigt sich mit Präferenzmodellen, welche direkt Rankings lernen und zurückgeben. In der Evaluation stellte sich heraus, dass der Regressionsansatz dem Präferenzmodell überlegen ist; das beste der getesteten Regressionsmodelle, basierend auf dem Random Forest Algorithmus, erreicht einen Wert von Kendalls Tau von bis zu 0,495 und Performanzverlustwerte von durchschnittlich nur 1,308 Prozentpunkten. Jedoch sind auch manche der getesten Präferenzmodelle in einigen Punkten wenigstens der zum Vergleich benutzten Baseline überlegen. Daher kann die Schlussfolgerung gezogen werden, dass wenigstens einige der benutzen Eigenschaften der Datensätze mit den Performanzwerten der Klassifizierungsalgorithmen zusammenhängen, und dieser Zusammenhang durch Regressionsmodelle sowohl als auch Präferenzmodelle ausgenutzt werden kann, wobei jedoch der Regressionsansatz letzterem im Rahmen der durchgeführten Evaluation überlegen ist.

>>>>

% Motivation
Die weltweit verfügbare Menge an Daten steigt immer weiter an, und doch bleibt der Anteil der tatsächlich analysierten Daten gering. Dieses Problem entsteht sehr wahrscheinlich, da nicht genug Experten verfügbar sind, um diese Analyse auszuführen, und es für Laien schwierig ist, das richtige Verfahren auszuwählen. Besonders im Bereich der Klassifikation gibt es viele Algorithmen, deren Performanz jedoch von Datensatz zu Datensatz variieren kann - es gibt also keinen einen besten Algorithmus, der immer eine gute Wahl darstellt. Eine Antwort auf diese Problem ist automatisiertes Machine Learning: Werkzeuge, für einen neuen Datensatz einen Lernalgorithmus, von dem sie schätzen dass dieser bezüglich dessen die besten Vorraussagen treffen kann, vorschlagen. Aktuelle Ansätze dieser Art führen diese Suche oft aus, ohne explizit zu berücksichtigen wie Eigenschaften des Datensatzes die Performanz von Klassifizierungsalgorithmen beeinflussen. Außerdem schlagen sie meist nur einen einzelnen Algorithmus vor, was den Nachteil hat, dass keine Alternativen zu dem vorgeschlagenen Algorithmus verfügbar sind, welche aufgrund der Ungenauigkeit von Vorhersagen möglicherweise eine gleicherwertige oder sogar bessere Vorhersagegenauigkeit auf diesem Datensatz haben. Daher werden in dieser Arbeit zwei verschiedene Ansätze, welche für einen neuen Datensatz ein Ranking von Klassifizierungsalgorithmen aufgrund von Eigenschaften des Datensatzes vorhersagt, vorgestellt und miteinander verglichen. Der erste der beiden Ansätze arbeitet auf Grundlage von Regressionsmodellen; ein Ranking wird hier konstruiert indem ein einzelnes Regressionsmodell für jeden Klassifizierungsalgorithmus einen Performanzwert vorhersagt, auf deren basierend dann das Ranking erstellt wird. Der zweite Ansatz beschäftigt sich mit Präferenzmodellen, welche direkt Rankings lernen und zurückgeben. In der Evaluation stellte sich heraus, dass der Regressionsansatz dem Präferenzmodell überlegen ist. Jedoch sind auch manche der getesten Präferenzmodelle in einigen Punkten wenigstens der zum Vergleich benutzten Baseline überlegen.

### 2018-02-17 Writing/Thesis/content/chapter-approach.tex
Some more corrections (writing)

In this chapter the two alternatives considered in this thesis to tackle the problem of ranking classification algorithms regarding properties of the data set. First, the theoretical approach is presented. Then, implementation details are discussed.

@@

\begin{table}[h]
\centering
	\begin{tabularx}{\textwidth}{X | X | X | X | X | X | X | X}
		%\hline
		MF 1			& MF 2		& ... 	& MF n		& PV 1 		& PV 2 		&	...	&	PV k 		\\ \hline
		$v_{11}$		& $v_{12}$	& ...	& $v_{1n}$	& $p_{11}$	& $p_{12}$	& 	...	&	$p_{1k}$		\\ 
		$v_{21}$		& $v_{22}$	& ...	& $v_{2n}$	& $p_{21}$	& $p_{22}$	& 	...	&	$p_{2k}$		\\ 
		...			& ...		& ...	& ...		& ...		& ...		&	...	&	...			\\ 
		$v_{m1}$		& $v_{m2}$	& ... 	& $v_{mn}$	& $p_{2m}$	& $p_{m2}$	& 	...	&	$p_{mk}$			 
	\end{tabularx}
	\label{tab:performanceValues}
	\caption{Training data for the rankers.}
\end{table}

\subsection{Regression Based Ranking}
% How is the regression ranking done
One possibility to derive a ranking of classifiers from this information when given a new data set is to use regression models. The idea is to use separate regression models to predict a performance value for each classifier, and to then derive an ordering from these predictions. This is done by splitting the data set compiled beforehand (see Tab. \ref{tab:performanceValues}) into separate data sets that each contain all meta features for all data sets, but only the performance values of one classifier, as illustrated in Table \ref{tab:regressionTables}. The target feature on these individual data sets, the performance value of the classifier, is a numeric value, and thus a regression model can be trained on each of them. All regression models therefore try to learn the target function metafeatures$\rightarrow$performance value for their respective classifier. Hence for a query instance, after computing the meta features, these are fed into each regression model and the predicted performance value for each classifier is saved. In a second step, the classifiers are ordered according to the predictions. This process the ranker passes through for making a prediction is presented in Figure \ref{fig:regression_ranker_model}.

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
	\begin{tabularx}{1.1\textwidth}{X | X | X | X}
		%\hline
		Meta Feature 1	& ... 	& Meta Feature n		& Performance Value 1	\\ \hline
		$v_{11}$			& ...	& $v_{1n}$			& $p_{11}$				\\ 
		$v_{21}$			& ...	& $v_{2n}$			& $p_{21}$				\\ 
		...				& ...	& ...				& ...					\\ 
		$v_{m1}$			& ... 	& $v_{mn}$			& $p_{m1}$				\\ \hline\hline
		Meta Feature 1	& ... 	& Meta Feature n		& Performance Value 2 	\\ \hline
		$v_{11}$			& ...	& $v_{1n}$			& $p_{12}$				\\ 
		$v_{21}$			& ...	& $v_{2n}$			& $p_{22}$				\\ 
		...				& ...	& ...				& ...					\\ 
		$v_{m1}$			& ... 	& $v_{mn}$			& $p_{m2}$					 
	\end{tabularx}
}	
\vspace{1em}
...
\vspace{1em}	
\resizebox{\textwidth}{!}{%
	\begin{tabularx}{1.1\textwidth}{X | X | X | X}
		%\hline
		Meta Feature 1	& ... 	& Meta Feature n		& Performance Value k 	\\ \hline
		$v_{11}$			& ...	& $v_{1n}$			& $p_{1k}$				\\ 
		$v_{21}$			& ...	& $v_{2n}$			& $p_{2k}$				\\ 
		...				& ...	& ...				& ...					\\ 
		$v_{m1}$			& ... 	& $v_{mn}$			& $p_{mk}$					 
	\end{tabularx}
}
	\caption{How the regression ranker processes the training data.}
	\label{tab:regressionTables}
\end{table}

@@

\subsection{Preference Based Ranking}
% How is the preference ranking done
The second possibility considered in this context is using preference learning to predict a ranking. Each instance of the data set generated beforehand (see Fig \ref{tab:performanceValues}) contains meta feature information for the considered data set and performance values of all classifiers. Similar to the regression alternative, the performance values can be converted to preference information by sorting the corresponding classifiers by their performance values for each data set. This leads to an ordering of classifiers associated with each instance, and implies that the preference learning task at hand is label ranking. In this case, there exists only one label ranking model which is fed the computed meta data for a new data set and returns an ordering of classifiers, which is depicted in Figure \ref{fig:preference_ranker_model}. Thereby it attempts to learn the mapping from meta features to an ordering of classifiers directly.

@@

\section{Implementation}
In general, all implementations for the approach are carried out in Java, particularly Java 8. However, some additional packages where used, which are pointed out briefly. For the implementation of the regression models, WEKA was used, a popular libary of machine learning algorithms that includes algorithms for classification, regression, and provides help in the evaluation of these learning algorithms \cite{hall2009weka}. Regarding the preference algorithms, the jPL framework was used \cite{intelligent2017jpl}. The jPL framework is a java framework for the evaluation of preference learning algorithms, it implements several tasks from the context of preference learning, including label ranking, and also provides a representation for data sets that contain label ranking information. The list of meta features used is taken from OpenML \cite{OpenML2013}, a website that not only provides a large number of openly accessible data sets, but also records performance values of learning algorithms on these data sets, and features properties of data sets. Thus the meta features where calculated with the help of an implementation provided by OpenML in its Evaluation Engine \cite{openMLEvaluationEngine}; The full list of meta features is included in Table \ref{tab:metaFeatureDetails} in the Appendix.

>>>>

In this chapter, the two alternatives considered in this thesis to tackle the problem of ranking classification algorithms regarding properties of a data set are explained. First, the theoretical approach is presented. Then, implementation details are discussed.

@@

\begin{table}[h]
\centering
	\begin{tabularx}{\textwidth}{X | X | X | X | X | X | X | X}
		%\hline
		MF 1			& MF 2		& ... 	& MF n		& PV 1 		& PV 2 		&	...	&	PV k 		\\ \hline
		$v_{11}$		& $v_{12}$	& ...	& $v_{1n}$	& $p_{11}$	& $p_{12}$	& 	...	&	$p_{1k}$		\\ 
		$v_{21}$		& $v_{22}$	& ...	& $v_{2n}$	& $p_{21}$	& $p_{22}$	& 	...	&	$p_{2k}$		\\ 
		...			& ...		& ...	& ...		& ...		& ...		&	...	&	...			\\ 
		$v_{m1}$		& $v_{m2}$	& ... 	& $v_{mn}$	& $p_{2m}$	& $p_{m2}$	& 	...	&	$p_{mk}$			 
	\end{tabularx}
	\caption{Training data for the rankers.}
	\label{tab:performanceValues}
\end{table}

\subsection{Regression Based Ranking}
% How is the regression ranking done
One possibility to derive a ranking of classifiers from this information when given a new data set is to use regression models. The idea is to use separate regression models to predict a performance value for each classifier, and to subsequently derive an ordering from these predictions. This is done by splitting the data set compiled beforehand (see Tab. \ref{tab:performanceValues}) into separate data sets that each contain all meta features for all data sets, but only the performance values of one classifier, as illustrated in Table \ref{tab:regressionTables}. The target feature on these individual data sets, the performance value of the classifier, is a numeric value, and thus, a regression model can be trained on each of them. All regression models therefore try to learn the target function \textit{meta features}$\rightarrow$\textit{performance value} for their respective classifier. Hence, for a query instance, after computing the meta features, these are fed into each regression model and the predicted performance value for each classifier is saved. In a second step, the classifiers are ordered according to the predictions. This generation of a ranking is presented in Figure \ref{fig:regression_ranker_model}.

\begin{table}[h]
\centering
\resizebox{\textwidth}{!}{%
	\begin{tabularx}{1.1\textwidth}{X | X | X | X}
		%\hline
		Meta Feature 1	& ... 	& Meta Feature n		& Performance Value 1	\\ \hline
		$v_{11}$			& ...	& $v_{1n}$			& $p_{11}$				\\ 
		$v_{21}$			& ...	& $v_{2n}$			& $p_{21}$				\\ 
		...				& ...	& ...				& ...					\\ 
		$v_{m1}$			& ... 	& $v_{mn}$			& $p_{m1}$				\\ \hline\hline
		Meta Feature 1	& ... 	& Meta Feature n		& Performance Value 2 	\\ \hline
		$v_{11}$			& ...	& $v_{1n}$			& $p_{12}$				\\ 
		$v_{21}$			& ...	& $v_{2n}$			& $p_{22}$				\\ 
		...				& ...	& ...				& ...					\\ 
		$v_{m1}$			& ... 	& $v_{mn}$			& $p_{m2}$					 
	\end{tabularx}
}	
\vspace{1em}
...
\vspace{1em}	
\resizebox{\textwidth}{!}{%
	\begin{tabularx}{1.1\textwidth}{X | X | X | X}
		%\hline
		Meta Feature 1	& ... 	& Meta Feature n		& Performance Value k 	\\ \hline
		$v_{11}$			& ...	& $v_{1n}$			& $p_{1k}$				\\ 
		$v_{21}$			& ...	& $v_{2n}$			& $p_{2k}$				\\ 
		...				& ...	& ...				& ...					\\ 
		$v_{m1}$			& ... 	& $v_{mn}$			& $p_{mk}$					 
	\end{tabularx}
}
	\label{tab:regressionTables}
	\caption{How the regression ranker processes the training data.}
\end{table}

@@

\subsection{Preference Based Ranking}
% How is the preference ranking done
The second possibility considered in this context is using preference learning to predict a ranking. Each instance of the data set generated beforehand (see Fig \ref{tab:performanceValues}) contains meta feature information for the considered data set and performance values of all classifiers. Similar to the regression alternative, the performance values can be converted to preference information by sorting the corresponding classifiers by their performance values for each data set. This leads to an ordering of classifiers associated with each instance, and implies that the preference learning task at hand is label ranking. In this case, there exists only one label ranking model which is fed the computed meta data for a new data set and returns an ordering of classifiers, which is depicted in Figure \ref{fig:preference_ranker_model}. Thereby, it attempts to learn the mapping from meta features to an ordering of classifiers directly.

@@

\section{Implementation}
In general, all implementations for the approach are carried out in Java, particularly Java 8. However, some additional packages were used, which are pointed out briefly. For the implementation of the regression models, WEKA was used, a popular libary of machine learning algorithms that includes algorithms for classification, regression, and provides help in the evaluation of these learning algorithms \cite{hall2009weka}. Regarding the preference algorithms, the jPL framework was used \cite{intelligent2017jpl}. The jPL framework is a Java framework for the evaluation of preference learning algorithms, it implements several tasks from the context of preference learning, including label ranking, and also provides a representation for data sets that contain label ranking information. The list of meta features used is taken from OpenML \cite{OpenML2013}, a website that not only provides a large number of openly accessible data sets, but also records performance values of learning algorithms on these data sets, and features properties of data sets. Thus the meta features were calculated with the help of an implementation provided by OpenML in its Evaluation Engine \cite{openMLEvaluationEngine}; The full list of meta features is included in Table \ref{tab:metaFeatureDetails} in the Appendix.

### 2018-02-17 Writing/Thesis/content/chapter-fundamentals.tex
Some more corrections (writing)

\begin{figure}
\centering
\includegraphics[width=\textwidth]{gfx/machine_learning.pdf}
\caption{The learning problem, adapted from \cite{abu2012learning}.}
\label{fig:learning_problem}
\end{figure}

@@

\subsection{Classification and Regression}
In addition to falling into the category of supervised learning, the problem is a classification problem. The nature of the prediction is to assign one of the three classes, or labels, 'iris setosa', iris virginica' and 'iris versicolor' to a new plant. Formally, this means that the output space $Y$ of the target function consists of a finite set of labels $\lbrace y_1,\dots,y_n\rbrace$, so that each instance of the data set is associated with a label $y_i \in Y$, in this case $Y=\lbrace 'iris-setosa','iris virginica', 'iris versicolor' \rbrace $.

Apart from classification, another important category of machine learning problems is regression problems. In the case of regression, the output space of the target function is no longer finite, but instead consist of real numbers. To reconsider the gardening example, the gardener could want to predict the height of a plant on the basis of the same features as used above. Likewise, they would have to observe a number of plants to gather feature values together with the expected target value, which instead of the species would then be the height of the plant in centimeters, that is $Y=\mathbb{R}$.

%TODO include a picture of the regression model mapping one of the width to height & say that this would be a linear model but others could be used

@@

\subsubsection{Ranking and Ordering}
To clarify the meaning of ranking and ordering, we return to the gardening example. After spending some time studying the different flowers, the gardener has realized he prefers certain iris species to others. He expresses his preference for a plant in a score, and labels all the plants in no particular order (see Fig. \ref{tab:ranking_vs_ordering}). An \textit{ordering} then is the result of simply ordering the labels in decreasing order of their respective score. It thus can also be given implicitly by just ordering the original classes, the flower names, as such being a permuation of the labels. A ranking, on the other hand, assigns a fixed position to each label; for instance the rank of 'iris virginica' will always be in the first position in the ranking, that is the number in that position denotes the \textit{rank} of the label, which is in turn defined as the label's position in the ordering. Therefore the values in the example in Table \ref{tab:ranking_vs_ordering} result in the ranking $[2,3,1]$, whereas the ordering is $[3,1,2]$. It is notable that these can be the same, and are easily convertible. However, in the remainder of this thesis mostly the term ranking is used to identify both ranking and ordering interchangeably, especially as usually only the term 'ranking' is used.

@@

\subsubsection{Label Ranking}
Similar to classification, in label ranking there is a finite set of labels $Y=\lbrace y_1,\dots,y_n \rbrace$, but it is not the output space. Instead, \citeauthor{DBLP:books/daglib/0025729} define the target function for label ranking as $f:X\rightarrow S_Y$ where the output space $S_Y$ contains all permutations over the set of labels $Y$ \cite{DBLP:books/daglib/0025729}. To come back to the topic of rankings and orderings, it becomes clear now that even though it is called label ranking, the labels are actually ordered. An example of label ranking could be, to stay in the domain of gardening, to learn the preferences of different gardeners towards iris plants. A label ranking data set used to learn this connection could, for example, map the favourite colour and height of a gardener to a ranking of the three iris plants.

@@

The approach of this thesis clearly falls under that category, as it is trying to learn whether a connection exists beween properties of a data set, that is its meta features, and the performance values of certain classification algorithms, and how well it can be exploited by regression and preference models respetively, if it exists. The benefit from meta learning in the conext of machine learning is that learning how properties of data sets influence the learning process of the classifiers could save much work; Often, much work is put into the fitting of an algorithm to a specific data set, but it is not transferred to other poblems that may be of the same nature.

@@

After having discussed the preliminaries that the theoretical approach of this thesis is built upon, how a data set can be split into training and test data to estimate the performance of learning algorithms and the measures that will be used in the evaluation of the different ranking methods are briefly reviewed. These include means of comparing the predicted rankings with actual rankings as well as a statistic that reveals if an advantage regarding one measure can be viewed as significant.

@@

To find out how well a classifier performs, it first needs to be decided which measure consider. There are a few possibilities for this, however in this thesis the term performance measure refers to the predictive accuracy of a classifier, that is the percentage of classes that a classifier correctly identified in a given data set. But since using the whole data set to train a classifier, that is build a model, would skew the analysis - the classifier has already 'seen' these examples - a method for dividing data sets into train and test sets is needed.

A popular approach is using a stratified split, for example splitting the data into 70\% test and 30\% training data, with the important aspect that both sub sets have equal (or at least very similar) ratios of classes [Reference Missing]. The learning algorithm is then trained with one part of the data (usually the larger), and its performance tested on the other. To get more stable results, often the process is repeated: The data is divided in k folds, k times, the learning algorithm is trained on k-1 folds, and tested on the left out subset. Results are averaged. This process is then called k-fold (stratified) cross validation [Reference Missing]. A variant of this is Monte Carlo crossvalidation, where after each time the algorithm has been trained and evaluated, a new split is computed [Reference Missing]. This especially causes the fact that the number of folds does not dictate the ratio of the split; one could for example have 5-folds Mone Carlo crossvalidation with a 70\%/30\% train/test split.

An alternative to using a stratified split is using a method called leave-one-out. This method essentially is, for a data set with n instances, n-fold crossvalidation (not stratified, naturally). This means that the algorithm is trained on n-1 instances and evaluated on the instance that has been left out.

@@

% Definition of Kendall
The Kendall Rank Correlation Coefficient, also known as Kendall's tau, is a measure of association between two rankings of the same items \cite{kendall1938new}. For two independent variables $X$ and $Y$ and and observations of values $(x_1,x_2,...,x_m)$ for $X$ and $(y_1,y_2,...,y_n)$ for $Y$ with unique $x_i$ and $y_i$ respectively, the coefficient is based on comparing pairs of observations. If for two pairs $(x_i,y_i)$ and $(x_j,y_j)$ both $x_i < x_j$ and $y_i < y_j$ (or the opposite) holds, they are viewed as concordant. If $x_i < x_j$ but $y_i > y_j$ (or the opposite), they are viewed as discordant. Other cases are not considered. The simplest version of the statistic, called $T_A$ is then defined as

$$T_A = \frac{n_c - n_d}{n_0},$$
with \\
$n_c$, the number of concordant pairs, \\
$n_d$, the number of discordant pairs, and \\
$n_0 = n * (n + 1) / 2$.

As we have seen in example above, can also have same rank for items. In order to account for this, a second statistic, $T_B$ is defined as

@@

\subsection{Loss}
While the Kendall rank correlation coefficient is an indication as to what degree a predicted ranking agrees with the correct ranking of classifiers for a certain data set, it does not convey the difference in performance experienced by the user when choosing the first of the recommended algorithms. This difference can be expressed in measure called loss \cite{DBLP:conf/mldm/LeiteBV12}, which is defined as the difference of the performance of the actual best algorithm and the actual performance of the first recommended algorithm.

@@

\subsection{Mann-Whitney U Test}
The measures previously presented can indicate if one of the ranking alternatives is superior to another regarding that measure, but one cannot conclude from those results alone whether this advantage is significant. The question here is whether the null-hypothesis that two distributions belong to the same population can be rejected. This can be determined with the help of the Mann-Whitney U, a non-parametric hypthesis test \cite{mann1947test}. The Mann-Whitney U is computed by, for two distributions one wants to analyze, counting how many times each value frome one distribution is lower than another value from the other distribution. If the U is very high (or low, depending on the hypothesis), the null hypothesis can be rejected. Whether this significance level is reached is expressed by the p-value: for a p-value below 0.05, The null hypothesis can be rejected. Therefore, if two distributions, for example regarding the loss values of two ranking approaches, suggest that one is superior to the other, and the p-value for the Mann-Whitney U is <= 0.05, this suggests that the difference is statistically significant.

>>>>

\begin{figure}
\centering
\includegraphics[width=\textwidth]{gfx/machine_learning.png}
\caption{The learning problem, adapted from \cite{abu2012learning}.}
\label{fig:learning_problem}
\end{figure}

@@

\subsection{Classification and Regression}
In addition to falling into the category of supervised learning, the problem is a classification problem. The nature of the prediction is to assign one of the three classes, also called labels, 'iris setosa', iris virginica' and 'iris versicolor' to a new plant. Formally, this means that the output space $Y$ of the target function consists of a finite set of labels $\lbrace y_1,\dots,y_n\rbrace$, so that each instance of the data set is associated with a label $y_i \in Y$, in this case $Y=\lbrace 'iris-setosa','iris virginica', 'iris versicolor' \rbrace $.

Apart from classification, another important category of machine learning problems are regression problems. In the case of regression, the output space of the target function is no longer finite, but instead consist of real numbers. To reconsider the gardening example, the gardener could want to predict the height of a plant on the basis of the same features as used above. Likewise, they would have to observe a number of plants to gather feature values together with the expected target value, which instead of the species would then be the height of the plant in centimeters, that is $Y=\mathbb{R}$.

@@

\subsubsection{Ranking and Ordering}
To clarify the meaning of ranking and ordering, we return to the gardening example. After spending some time studying the different flowers, the gardener has realized he prefers certain iris species to others. He expresses his preference for a plant in a score, and labels all the plants in no particular order (see Fig. \ref{tab:ranking_vs_ordering}). An \textit{ordering} then is the result of simply ordering the labels in decreasing order of their respective score. It thus can also be given implicitly by just ordering the original classes, the flower names, as such being a permuation of the labels. A ranking, on the other hand, assigns a fixed position to each label; for instance the rank of 'iris virginica' will always be in the first position in the ranking, that is the number in that position denotes the \textit{rank} of the label, which is in turn defined as the label's position in the ordering. Therefore, the values in the example in Table \ref{tab:ranking_vs_ordering} result in the ranking $[2,3,1]$, whereas the ordering is $[3,1,2]$. It is notable that these can be the same, and are easily convertible. However, in the remainder of this thesis mostly the term ranking is used to identify both ranking and ordering interchangeably, especially as usually only the term 'ranking' is used.

@@

\subsubsection{Label Ranking}
Similar to classification, in label ranking there is a finite set of labels $Y=\lbrace y_1,\dots,y_n \rbrace$, but it is not the output space. Instead, \citeauthor{DBLP:books/daglib/0025729} define the target function for label ranking as $f:X\rightarrow S_Y$ where the output space $S_Y$ contains all permutations over the set of labels $Y$ \cite{DBLP:books/daglib/0025729}. To come back to the topic of rankings and orderings, it becomes clear now that even though it is called label ranking, the labels are actually ordered. An example of label ranking could be, to stay in the domain of gardening, to learn the preferences of different gardeners towards iris plants. A label ranking data set used to learn this connection could, for example, map the favorite color and height of a gardener to a ranking of the three iris plants.

@@

The approach of this thesis clearly falls under that category, as it is trying to learn whether a connection exists between properties of a data set, that is its meta features, and the performance values of certain classification algorithms, and how well it can be exploited by regression and preference models respectively, if it exists. The benefit from meta learning in the context of machine learning is that learning how properties of data sets influence the learning process of the classifiers could save much work; often, much work is put into the fitting of an algorithm to a specific data set, but it is not transferred to other problems that may be of the same nature.

@@

After having discussed the preliminaries that the theoretical approach of this thesis is built upon, how a data set can be split into training and test data to estimate the performance of learning algorithms and the measures that will be used in the evaluation of the different ranking methods are briefly reviewed. These include means of comparing the predicted rankings with actual rankings as well as a statistic that reveals whether an advantage regarding one measure can be viewed as significant.

@@

To find out how well a classifier performs, it first needs to be decided which measure to consider. There are a few possibilities for this, however in this thesis the term performance measure refers to the predictive accuracy of a classifier, that is the percentage of classes that a classifier correctly identified in a given data set. But since using the whole data set to train a classifier, that is build a model, would skew the analysis - the classifier has already 'seen' these examples - a method for dividing data sets into train and test sets is needed.

A popular approach is using a stratified split, for example splitting the data into 70\% training and 30\% test data, with the important aspect that both subsets have equal (or at least very similar) ratios of classes [Reference Missing]. The learning algorithm is then trained with one part of the data (usually the larger), and its performance tested on the other. To get more stable results, often the process is repeated: The data is divided in k folds, k times, the learning algorithm is trained on k-1 folds, and tested on the left out subset. Results are averaged. This process is then called k-fold (stratified) cross-validation [Reference Missing]. A variant of this is Monte Carlo cross-validation, where after each time the algorithm has been trained and evaluated, a new split is randomly computed [Reference Missing]. This especially causes the fact that the number of folds does not dictate the ratio of the split; one could for example have 5-folds Monte Carlo cross-validation with a 70\%/30\% train/test split.

An alternative to using a stratified split is using a method called leave-one-out. This method essentially is, for a data set with n instances, n-fold cross-validation (not stratified, naturally). This means that the algorithm is trained on n-1 instances and evaluated on the instance that has been left out.

@@

% Definition of Kendall
The Kendall Rank Correlation Coefficient, also known as Kendall's tau, is a measure of association between two rankings of the same items \cite{kendall1938new}. For two independent variables $X$ and $Y$ and observations of values $(x_1,x_2,...,x_m)$ for $X$ and $(y_1,y_2,...,y_n)$ for $Y$ with unique $x_i$ and $y_i$ respectively, the coefficient is based on comparing pairs of observations. If for two pairs $(x_i,y_i)$ and $(x_j,y_j)$ both $x_i < x_j$ and $y_i < y_j$ (or the opposite) holds, they are viewed as concordant. If $x_i < x_j$ but $y_i > y_j$ (or the opposite), they are viewed as discordant. Other cases are not considered. The simplest version of the statistic, called $T_A$ is then defined as

$$T_A = \frac{n_c - n_d}{n_0},$$
with \\
$n_c$, the number of concordant pairs, \\
$n_d$, the number of discordant pairs, and \\
$n_0 = n * (n - 1) / 2$.

To account for ties in the rankings, a second statistic, $T_B$ is defined as

@@

\subsection{Loss}
While the Kendall rank correlation coefficient is an indication as to what degree a predicted ranking agrees with the correct ranking of classifiers for a certain data set, it does not convey the difference in performance experienced by the user when choosing the first of the recommended algorithms. This difference can be expressed in a measure called loss \cite{DBLP:conf/mldm/LeiteBV12}, which is defined as the difference of the performance of the actual best algorithm and the actual performance of the first recommended algorithm.

@@

\subsection{Mann-Whitney U Test}
The measures previously presented can indicate if one of the ranking alternatives is superior to another regarding that measure, but one cannot conclude from those results alone whether this advantage is significant. The question here is whether the null-hypothesis that two distributions belong to the same population can be rejected. This can be determined with the help of the Mann-Whitney U, a non-parametric hypthesis test \cite{mann1947test}. The Mann-Whitney U is computed by, for two distributions one wants to analyze, counting how many times each value from one distribution is lower than another value from the other distribution. If the U is very high (or low, depending on the hypothesis), the null hypothesis can be rejected. Whether this significance level is reached is expressed by the p-value: for a p-value below 0.05, the null hypothesis can be rejected. Therefore, if two distributions, for example regarding the loss values of two ranking approaches, suggest that one is superior to the other, and the p-value for the Mann-Whitney U is <= 0.05, this suggests that the difference is statistically significant.
